Reanalysis Step 3: Propensity Score Analysis with Twang

### 2025-08-26 reanalysis-2-multiple_imputation.qmd
better plots

```{r imputation-diagnostics}
# Check imputed values for key variables

# CCI - compare observed vs imputed
if(sum(is.na(df_mice$CCI)) > 0) {
  densityplot(mice_results, ~ CCI) |> print()
  stripplot(mice_results, CCI ~ .imp, pch = 20, cex = 1.2) |> print()
}

# DCSI - compare observed vs imputed  
if(sum(is.na(df_mice$DCSI)) > 0) {
  densityplot(mice_results, ~ DCSI) |> print()
  stripplot(mice_results, DCSI ~ .imp, pch = 20, cex = 1.2) |> print()
}

# Baseline VA - compare observed vs imputed
if(sum(is.na(df_mice$baseline_VA_logMAR)) > 0) {
  densityplot(mice_results, ~ baseline_VA_logMAR) |> print()
  stripplot(mice_results, baseline_VA_logMAR ~ .imp, pch = 20, cex = 1.2) |> print()
}

@@

```{r summary}
summary_table <- tibble(
  Metric = c(
    "Original observations",
    "Variables in imputation model",
    "Variables with missing data",
    "Number of imputations",
    "Maximum missing percentage",
    "Imputation iterations"
  ),
  Value = c(
    nrow(df_mice),
    ncol(df_mice),
    sum(mice_results$nmis > 0),
    n_imputations,
    paste0(round(max(mice_results$nmis/nrow(df_mice)) * 100, 1), "%"),
    mice_results$iteration
  )
)

>>>>

**Note:** In the stripplots below, the leftmost column (labeled 1) shows the **original observed data** with missing values. Columns 2 through `r n_imputations + 1` show the imputed datasets.

```{r imputation-diagnostics}
# Check imputed values for key variables
# In mice stripplots: .imp = 1 is original data, .imp = 2+ are imputations

# CCI - compare observed vs imputed
if(sum(is.na(df_mice$CCI)) > 0) {
  cat("### Charlson Comorbidity Index (CCI)\n")
  cat("**Density plot:** Blue = observed data, Red = imputed values\n")
  densityplot(mice_results, ~ CCI) |> print()

  cat("\n**Stripplot:** Column 1 = original data (with gaps for missing), Columns 2+ = imputed datasets\n")
  stripplot(mice_results, CCI ~ .imp, pch = 20, cex = 1.2,
            xlab = "Dataset (1=Original, 2+=Imputations)",
            main = "CCI Distribution: Original vs Imputed") |> print()
  cat("\n")
}

# DCSI - compare observed vs imputed  
if(sum(is.na(df_mice$DCSI)) > 0) {
  cat("### Diabetes Complications Severity Index (DCSI)\n")
  cat("**Density plot:** Blue = observed data, Red = imputed values\n")
  densityplot(mice_results, ~ DCSI) |> print()

  cat("\n**Stripplot:** Column 1 = original data (with gaps for missing), Columns 2+ = imputed datasets\n")
  stripplot(mice_results, DCSI ~ .imp, pch = 20, cex = 1.2,
            xlab = "Dataset (1=Original, 2+=Imputations)",
            main = "DCSI Distribution: Original vs Imputed") |> print()
  cat("\n")
}

# Baseline VA - compare observed vs imputed
if(sum(is.na(df_mice$baseline_VA_logMAR)) > 0) {
  cat("### Baseline Visual Acuity (logMAR)\n")
  cat("**Density plot:** Blue = observed data, Red = imputed values\n")
  densityplot(mice_results, ~ baseline_VA_logMAR) |> print()

  cat("\n**Stripplot:** Column 1 = original data (with gaps for missing), Columns 2+ = imputed datasets\n")
  stripplot(mice_results, baseline_VA_logMAR ~ .imp, pch = 20, cex = 1.2,
            xlab = "Dataset (1=Original, 2+=Imputations)",
            main = "Baseline VA Distribution: Original vs Imputed") |> print()
  cat("\n")
}

@@

```{r summary}
# Count variables more precisely
# Total columns in df_mice includes row_id which we added
total_vars <- ncol(df_mice)
id_vars <- 2  # row_id and e_mrn_deidentified
substantive_vars <- total_vars - id_vars  # Variables actually used in the imputation model

summary_table <- tibble(
  Metric = c(
    "Original observations",
    "Total variables in dataset",
    "ID variables (excluded from model)",
    "Substantive variables in imputation model",
    "Variables with missing data",
    "Variables actually imputed",
    "Number of imputations",
    "Maximum missing percentage",
    "Imputation iterations"
  ),
  Value = c(
    nrow(df_mice),
    total_vars,
    id_vars,
    substantive_vars,
    sum(mice_results$nmis > 0),
    length(vars_to_impute),
    n_imputations,
    paste0(round(max(mice_results$nmis/nrow(df_mice)) * 100, 1), "%"),
    mice_results$iteration
  )
)

### 2025-08-26 reanalysis-3-propensity_scores.qmd
better plots

  cat("#### Balance Summary Table\n\n")
  summary(twang_model) |> print() |> kable()
  cat("\n\n")
}
```

>>>>

  cat("#### Balance Summary Table\n\n")

  # Get balance statistics from the twang model
  # The bal.table function gives us the detailed balance statistics
  bal_table <- bal.table(twang_model, digits = 4)

  # Extract the summary statistics for the optimal stopping point
  # twang stores these in the desc slot
  es_mean_stats <- twang_model$desc$es.mean

  # Create a clean summary table
  summary_df <- data.frame(
    Metric = c("N Treated", "N Control", 
               "ESS Treated", "ESS Control",
               "Max ES", "Mean ES", "Max KS",
               "Iterations"),
    Unweighted = c(
      es_mean_stats$n.treat,
      es_mean_stats$n.ctrl,
      es_mean_stats$n.treat,  # ESS = n for unweighted
      es_mean_stats$n.ctrl,    # ESS = n for unweighted
      round(bal_table$unw$max.es, 4),
      round(bal_table$unw$mean.es, 4),
      round(bal_table$unw$max.ks, 4),
      NA
    ),
    Weighted = c(
      es_mean_stats$n.treat,
      es_mean_stats$n.ctrl,
      round(es_mean_stats$ess.treat, 2),
      round(es_mean_stats$ess.ctrl, 2),
      round(bal_table$es.mean$max.es, 4),
      round(bal_table$es.mean$mean.es, 4),
      round(bal_table$es.mean$max.ks, 4),
      es_mean_stats$n.trees
    )
  )

  kable(summary_df, 
        caption = paste("Balance Summary for Imputation", i),
        align = c('l', 'r', 'r'))
  cat("\n\n")
}
```
